Potential Factors of Decreasing Birth Rates in the US since the Great Recession and Beyond

### 2024-02-11 outputs/paper/paper.qmd
scatter plot
update

## Trend in US Birth Rates
The overall trend fluctuates from time to time but the highest births moment is recorded as in 1991 for the rate of 71

@@

@@

In the discussion section, and any other relevant section, please be sure to discuss ethics and bias, with reference to relevant literature.

>>>>

## Trend in US Birth Rates
The overall trend fluctuates over time, with the highest number of births recorded in 1990 at a rate of 70.9. The second-highest peaks occurred in 1991 and 2007, with rates of 69.3. These peaks are accompanied by valleys. However, the difference between the birth rates in 1980 and 2020 is 12.6, and the gap between the maximum and minimum points during this period is 15.1. These calculations clearly demonstrate the variation over time and suggest that there may be factors influencing the gradual negative slope of the trend. Figure 2 presents the scatter plot version of Figure 1, providing a clearer representation of the negative trend observed in the line graph.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-scatter-graph
#| fig-cap: "Negative Trend in US Birth Rates"

# Plotting the data with scatter plot and trend line
ggplot(data, aes(x = year, y = brate_all)) +
  geom_point(color = "skyblue") +
  geom_smooth(method = "lm", se = FALSE, color = "pink") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  annotate("text", x = 2009, y = 70, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
  labs(x = "Year", y = "Births per 1,000 women age 15-44",
       caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
  theme_classic()
```

@@
